\LARGE \bf A Hybrid Control Framework for Autonomous Vehicles at Uncontrolled Intersections

### 2019-01-22 conference/iv2019/manuscript.tex
making edits from Vijay

\author{Nitin R. Kapania$^{1}$, Francesco Borrelli, J Christian Gerdes% <-this % stops a space
\thanks{*This work was supported by the Hyundai Center of Excellence}% <-this % stops a space
\thanks{$^{1}$Nitin Kapania is with the Dept. of Mechanical Engineering, Stanford University, and is a visiting scholar at UC Berkeley}%
}

@@

As autonomous vehicles (AVs) inch closer to reality, a central requirement for acceptance will be earning the trust of humans in everyday driving situations. In particular, the interaction between AVs and pedestrians is of
high importance, as every human is a pedestrian at some point of the day. This paper considers the interaction of a pedestrian and an autonomous vehicle at a mid-block, uncontrolled intersection where there is ambiguity
over when the pedestrian should cross and when and how the vehicle should yield. By modeling pedestrian behavior through the concept of gap acceptance, the authors show that a hybrid controller with just four distinct modes allows an autonomous vehicle to successfully interact with a pedestrian across a continuous spectrum of possible crosswalk-entry behaviors. The controller is validated in simulation through comparison with a previously published control policy obtained through solution of a POMDP, and experimental results are provided on a Hyundai Genesis vehicle for a virtual pedestrian.

@@

While autonomous vehicles have the potential to save thousands of lives every year and create tremendous societal benefits \cite{Fagnant2015}, widespread adoption is unlikely until AVs gain the broad trust of society. Given that every human is a pedestrian at some point during the day, one of the central ways that autonomous vehicles will be evaluated is through their interactions with pedestrians. In fact, recent events with industry leader Waymo highlight that AV-pedestrian interaction has still not been perfected\cite{MattRichtelandConorDougherty2015}. In general, interactions with pedestrians are complex, even for experienced human drivers. Issues such as lack of visibility, improper communication, poorly marked roads, and distraction on both the driver or pedestrian side can lead to accidents and fatalities. From 2015-2016, pedestrian fatalaties increased by 9\% to 5987, representing the highest number since 1990, and also representing 16\% of all automotive fatalities \cite{HighwayTrafficSafetyAdministration2016}. As autonomous vehicles inch closer to widespread adoption, they must have a clear control strategy for pedestrian interaction that can handle a wide variety of pedestrian behaviors while maintaining a reasonable flow of traffic.

@@

The need to further understand pedestrian behavior for autonomous driving has created a growing body of interdisciplinary research. One branch of research has focused around modeling pedestrian behavior given various sensor inputs. Keller et al \cite{Keller2014} presented a study on pedestrian path prediction and action classification (e.g. crossing vs. waiting) using Gaussian process dynamical models and trajectory matching from optical data. Several other techniques for pedestrian trajectory prediction have been proposed, including LQR \cite{Batkovic}, set-based reachability analysis \cite{Koschi2018}, and Markov processes \cite{Karasev2016}

Another branch of literature is focused on developing predictive models for pedestrian crossing. While typically studied for purposes of road design, this body of literature holds promising insights for AV designers. For example, Schroeder and Rouphail \cite{Schroeder2011} explored factors associated with driver yielding behavior at unsignalized pedestrian crossings. Using logistic regression, the authors found that drivers are more likely to yield to assertive pedestrians who walk briskly in their approach to a crosswalk. Kadali and Perumal \cite{RaghuramKadali2012} studied the ``gap acceptance" behavior of pedestrians at mid-block crosswalks through a video graphic survey, and found that the gap accepted for crossing was explained by factors such as crossing direction, vehicle speed, and pedestrian age. Yannis et al. \cite{Yannis2013} and \cite{Sun2002} also found that gap acceptance was influenced by the size of the oncoming vehicle and the presence of other pedestrians.  Lee and Aty \cite{Lee2005} studied interactions in the form of crashes, and found crashes were linked with higher daily traffic.

Finally, a small but rapidly growing body of literature specifically studies the interaction between pedestrians and autonomous vehicles at crosswalks. An excellent review of these studies was conducted by \cite{Rasouli}. As an example, Rothenbucher et al. \cite{Rothenbucher2016} studied the interaction between pedestrians and driverless vehicles by constructing a car seat costume to disguise a driver. The authors noted that pedestrians overall managed interactions at crosswalks effectively, but later mentioned increased uncertainty about the autonomous vehicles behavior. To improve the issue of trust, several researchers have developed external interfaces to more clearly broadcast the intent of the autonomous vehicle \cite{Matthews}, \cite{Lagstrom2015}.

Several real-world studies of AV-pedestrian interaction note that once a local population learned a vehicle was programmed to be perfectly safe, pedestrians would regularly take advantage of the AV and walk in front of it. \cite{Madigan}. This illustrates an issue with automated driving in that an overly conservative crossing algorithm will often be taken advantage of or cause confusion among pedestrians. Camara et al \cite{Camara2018} attempted to model the natural negotation for priority between a pedestrian and an AV at an intersection using the framework of ``sequential chicken" adopted from game theory. Chen et al. \cite{Chen} notes the need for a tradeoff between passive and aggressive driving behavior, and developed a stochastic model of pedestrian behavior to evaluate proposed AV control policies. Control polices for pedestrian interaction are also developed by \cite{Bandyopadhyay}, who proposes a Mixed-observable Markov Decision Process (MOMDP) to incorporate the intention uncertainty of a pedestrian. A POMDP formulation is also proposed by Thornton \cite{Thornton2018}, who proposes navigation of the pedestrian uncertainty through value-sensitive design.

@@

This paper aims to address this gap by developing a hybrid control architecture that accounts for several distinct pedestrian modes of behavior at an unsignalized crosswalk. An unsignalized crosswalk is chosen as it is the trickiest crosswalk for a pedestrian and vehicle to navigate, as there is no explicit declaration of whose turn it is cross. Simulations show that the state machine controller is able to handle a continuous spectrum of pedestrian \textit{gap acceptance} behavior, tolerating a range of highly conservative to highly aggressive pedestrians. Moreover, this is possible with just four distinct states. Simulations are also used to compare the closed-loop controller with the controller developed by \cite{Thornton2018}. Finally, experimental results are shown on a real vehicle to demonstrate the feasibility of the proposed state machine controller.

@@

In general, pedestrian crosswalks may be controlled or uncontrolled. In the case of the former, control devices such as stop signs or walk signals guide the interaction of vehicles and pedestrians explicitly. Fig.~\ref{fig:schematic} shows the latter example, in which a pedestrian must select a gap in traffic flow and cross.

\begin{figure}
\centering
\includegraphics[width=3.5in]{figures/example.eps}
\caption{Schematic of pedestrian approaching a stream of traffic at a mid-block intersection.}
\label{fig:schematic}
\end{figure}

In general, right-of-way for uncontrolled intersections is complex. For example, nine states and the District of Columbia require motorists to stop when approaching a pedestrian in an uncontrolled crosswalk. Six states require a motorist to stop when a pedestrian is upon the same half of the roadway or within one lane of the lane that the motorist is traveling upon. Another nineteen states require a motorist to yield when a pedestrian is upon any portion of the roadway, and another 20 states mandate that motorists yield when a pedestrian is upon the same half of the roadway or approaching closely from the opposite side of the roadway \cite{NCSL}.

@@

Similar to time to collision (TTC), the safety gap is a measure of how much time there is before the vehicle would enter the crosswalk if it kept it's current speed constant. When faced with a stream of traffic at a crosswalk, a pedestrian inherently decides how much of a gap to accept before crossing. The average gap acceptance is reported in the literature to be between 3-7 seconds, meaning pedestrians usually do not cross if the vehicle would enter the crosswalk in under three seconds \cite{DiPietroCharlesMandKing1970}, and are very likely to cross when they have more than seven seconds \cite{Schmidt2009}.

\subsection{Problem Formalization and System Requirements}
\label{sec:probform}

A diagram of the relevant state variables is shown below in Fig.~\ref{fig:diagram1}. We assume that knowledge of distance $d$ to a fixed stopping point 3-5 meters ahead of the crosswalk can be estimated by a perception system, and that the velocity of the vehicle $\dot{d}$ is known as well. Furthermore, we also assume the perception system is able to determine the position of the pedestrian $x_p$ within the crosswalk, and is able to determine a velocity estimate $\dot{x_p}$ for the pedestrian as well.

\begin{figure}
\centering
\includegraphics[width=1in]{figures/diagram.eps}
\caption{Description of relevant problem variables.}
\label{fig:diagram1}
\end{figure}

In general, this control problem is difficult as there are multiple stakeholders and therefore multiple competing objectives that must be considered. For example, relevant stakeholders for this problem are the autonomous vehicle and its passengers, the pedestrian(s), and the other vehicles in the traffic stream.

As a result, an algorithm that simply yields every time a pedestrian is near the crosswalk is likely to be too conservative, as there is a likelihood the pedestrian is waiting for a larger gap in the traffic stream to cross or is walking through the sidewalk without intending to cross in the first place. To formalize the system requirements for this multi-stake holder problem, we propose the following \textit{design constraints}, extended from the engineering specifications proposed by Thornton \cite{Thornton2018}.

\begin{center}
\begin{tabular}{ l| l }
\label{tb:specs}
 \textbf{Engineering} & \textbf{Design} \\
 \textbf{Specification} & \textbf{Constraint}\\\hline 
 1. Safety and Legality & a. Avoid collisions for all\\
                     & possible accepted gaps                        \\
                     &b. Ability to follow stop/yield    \\
                     &laws for all 50 U.S. states\\

               & \\
 3. Smoothness & a. Low accel: Nominal brake / \\ 
 			   & accel limits of 2 m/s/s \\
 			   & b. Stop within 3-5 meters\\
 			   & of crosswalk

@@

The proposed control architecture is shown below in Fig.~\ref{fig:hybridController}, and the algorithm for operation is shown in Algorithm \ref{alg:hybridController}. At heart, the control algorithm uses a feedback-feedforward methodology to compute desired acceleration commands $\ddot{d}$, but computes these commands differently depending on which of four discrete states the controller is in. The next sections will describe the four states.

\begin{figure}
\centering
\includegraphics[width=3.5in]{figures/stateDiagram.eps}
\caption{Schematic of hybrid control structure.}
\label{fig:hybridController}
\end{figure}

@@

		\ElsIf{$d > \frac{\dot{d}^2}{2a_\mathrm{cmfrt}}$}
		\State $state \leftarrow YIELDING$

	\ElsIf{$d < \frac{\dot{d}^2}{2a_\mathrm{cmfrt}} \hspace{1mm} \mathbf{and} \hspace{1mm} d > \frac{\dot{d}^2}{2a_\mathrm{max}}$}
	\State $state \leftarrow HARD\hspace{1mm}BRAKING$

@@

	\EndIf
	\EndIf
	\EndIf
\vspace{3mm}
\If {$state = YIELDING$}
\If {$d > \frac{\dot{d}^2}{2a_\mathrm{cmfrt}} + t_{delay}\dot{d}$}
	\State $\ddot{d}_{des} = 0$
	\State $\dot{d}_{des} = v_\mathrm{speed limit}$
\Else 
	\State $\ddot{d}_{des} = -a_\mathrm{cmf}$
	\State $\dot{d}_{des} = \mathrm{getDesiredSpeed}(d, \dot{d})$
\EndIf 
	\State $\ddot{d} = \ddot{d}_{des} + k_s(\dot{d} - \dot{d}_{des})$

@@

\EndIf
\If {$state = SPEED\hspace{1mm}UP$}
	\State $\ddot{d} = a_\mathrm{cmf}$ 	
	\If {$\textbf{not} \hspace{1mm} \mathrm{inCrosswalk}(x_p)$} \State $state \leftarrow DRIVING$

@@

\subsection{Nominal Driving}
In the nominal state, the vehicle attempts to drive through the crosswalk at the speed limit $v_\mathrm{speedlimit}$. A simple proportional speed control is applied to keep the vehicle at the speed limit, achieving design constraint 2.a.:

@@

The controller remains in the driving mode unless the pedestrian begins to enter the crosswalk while $d > 0$. To allow more time for the controller to respond, we allow the $\mathrm{inCrosswalk}(x_p, \dot{x}_p)$ indicator function to be triggered the moment the pedestrian approaches the crosswalk from the sidewalk - e.g. once the pedestrian velocity $\dot{x}_p$ becomes non-zero.

 Note that according to design constraint 1.b, the definition of entering the crosswalk will vary according to which of the fifty US states the vehicle is in. For example, in California, the pedestrian entering any portion of the crosswalk should trigger the control mode to change - while in Louisiana, the pedestrian entering the same half of the roadway as the vehicle should trigger the change \cite{NCSL2018}.

@@

Frequently he vehicle will pass through the intersection well before the pedestrian will reach the vehicle's lane. Assuming the vehicle is operating in a US state where stopping is not explicitly required, design constraint 2.a suggests it is desirable for the vehicle to continue through the intersection.

This can be formalized by calculating the \textit{time advantage} $t_adv$, also known as the \textit{time to collision} as follows:

\begin{equation}
t_{adv} = \frac{d}{\dot{d}} - \frac{x_v - x_p}{\dot{x_p}}
\end{equation}

Where $x_v$ is the $x$ position of the vehicle in the crosswalk. The control algorithm in this work will explicitly allow the vehicle to continue driving through the intersection as long as $t_{adv}$ exceeds a specfiied threshold $t_{max}$.

@@

If the time advantage is not sufficient for the vehicle to pass through, the autonomous vehicle must slow down. The preferred option to meet design constraint 3.a is to brake at a low, comfortable deceleration $-a_{cmf}$ for the passenger. Note that design constraint 3.b is satisfied by having $d=0$ be defined to be a point several meters ahead of the crosswalk.

Comfortable braking at a deceleration $-a_{cmf}$ is possible if sufficient braking distance exists. The braking distance is derived kinematically as

\begin{equation}
d_{brake} = \frac{\dot{d}^2}{2a_{cmf}}
\end{equation}

If this holds true when the pedestrian starts to enter the crosswalk, the car enters a $Yielding$ state.

@@

In the unlikely event a pedestrian begins to cross when the time gap is low (e.g. under 3 seconds), the vehicle will need to prioritize design constraint 1.a and come to a stop at a deceleration higher than $-a_{cmf}$. This occurs when the following conditions holds:

\begin{equation}
\frac{\dot{d}}{2a_{max}} < d < \frac{\dot{d}^2}{2a_{cmf}}
\end{equation}

@@

\subsection{Speeding Up}

As a final condition, consider the case where the pedestrian enters the crosswalk just as the vehicle is crossing $d = 0$ - in this case, it makes little sense to decelerate, as the vehicle has insufficient space to stop before the crosswalk and will risk being rear-ended (Design Constraint 2.a) or stopping in the crosswalk. It makes more sense in this condition for the vehicle to speed up and exit the crosswalk quickly. As a result, if $d < \frac{\dot{d}}{2a_{max}}$, the vehicle speeds up.

@@

In the yielding state, the vehicle follows a different version of the feedback-feedforward dynamics. The vehicle first follows the speed limit until it reaches the critical value of $\frac{\dot{d}^2}{2a_{cmf}}$. In practice, an additional term $t_delay\dot{d}$ is added to compensate for the delay in brake communication.

At the critical istance, the vehicle decelerates at the uniform yielding deceleration of $-a_{cmf}$, with an additional feedback term $k_s(\dot{d} - \dot{d}_des$), where $\dot{d}_des$ is the desired velocity. The feedback term helps bring the vehicle speed to 0 at the desired stopping point.

The desired velocity profile $\dot{d}(d)$ for a vehicle decelerating at constant acceleration is given via integration:

\begin{equation}
\dot{d}(d) = \sqrt{2a_{cmf}(d-d_o) + \dot{d}_o^2}
\end{equation}

Where $d_o$ and $\dot{d}_o$ is the value of $d$ and $\dot{d}$ when the controller first enters the yielding state. Simple inspection shows that the desired velocity is $\dot{d}_o$ when $d = d_o$ and is 0 when $d = 0$ (recall that we enter the yielding state when $d_o = \frac{\dot{d_o}^2}{2a_{cmf}}$.

The controller exits the yielding mode and returns to the driving mode once the pedestrian clears the crosswalk.

@@

In the hard braking state, the vehicle follows a similar feedback-feedforward algorithm, but the desired feedforward deceleration in this case is given by $\ddot{d} = -\frac{d^2}{2d}$. Integrating to find the desired speed profile $\dot{d}(d)$ yields the following deceleration:

\begin{equation}
\dot{d}(d) = \frac{\dot{d}_o}{\sqrt{d_o}}\sqrt{d}
\end{equation}

Where $d_o$ and $\dot{d}_o$ is the value of $d$ and $\dot{d}$ when the controller first enters the braking state. Again, the controller exits the braking mode and returns to the driving mode once the pedestrian clears the crosswalk.

@@

In the speed up state, there is no need to follow an exact speed profile as the vehicle is merely trying to exit the crosswalk area slightly faster. In this case, the commanded acceleration $\ddot{d} = a_{cmf}$ until the pedestrian clears the crosswalk.

@@

In order to check performance of the algorithm, both simulation and experimental studies were undertaken. The simulation results focused on testing Algorithm 1 against the design requirements from Table \ref{tb:specs} over many pedestrian crossings.

The algorithm is validated in simulation via a built-from-scratch crosswalk simulation environment (Fig. \ref{fig:simFramework}) developed in Python. The simulation environment is built on the ROS \cite{ROS} robotics interface to enable easy porting of the control architecture to the experimental vehicle in Section \ref{sec:expres}. The simulation includes features such as multiple lanes, the pedestrian crossing on the right or the left, and animation capabilities.

>>>>

\author{Nitin R. Kapania$^{1}$, Vijay Govindarajan$^{2}$, Francesco Borrelli$^{2}$, J Christian Gerdes$^{1}$% <-this % stops a space
\thanks{*This work was supported by the Hyundai Center of Excellence}% <-this % stops a space
\thanks{$^{1}$Dept. of Mechanical Engineering, Stanford University. Nitin Kapania is also a visiting scholar at UC Berkeley.}%
\thanks{$^{2}$Dept. of Mechanical Engineering, UC Berkeley.}%
}

% \thanks{$^{2}$Dept. of Mechanical Engineering, UC Berkeley. 
% }

@@

As autonomous vehicles (AVs) inch closer to reality, a central requirement for acceptance will be earning the trust of humans in everyday driving situations. In particular, the interaction between AVs and pedestrians is of
high importance, as every human is a pedestrian at some point of the day. This paper considers the interaction of a pedestrian and an autonomous vehicle at a mid-block, uncontrolled intersection where there is ambiguity
over when the pedestrian should cross and when and how the vehicle should yield. By modeling pedestrian behavior through the concept of gap acceptance, the authors show that a hybrid controller with just four distinct modes allows an autonomous vehicle to successfully interact with a pedestrian across a continuous spectrum of possible crosswalk entry behaviors. The controller is validated through extensive simulation, and experimental results are provided on a Hyundai Genesis vehicle for a virtual pedestrian.

@@

While autonomous vehicles have the potential to save thousands of lives every year and create significant societal benefits \cite{Fagnant2015}, widespread adoption is unlikely until AVs gain the broad trust of society. Given that every human is a pedestrian at some point during the day, one of the central ways that autonomous vehicles will be evaluated is through their interactions with pedestrians. In fact, recent events with industry leader Waymo highlight that AV-pedestrian interaction has still not been perfected \cite{MattRichtelandConorDougherty2015}. In general, interactions with pedestrians are complex, even for experienced human drivers. Issues such as lack of visibility, improper communication, poorly marked roads, and distraction on both the driver or pedestrian side can lead to accidents and fatalities. From 2015-2016, pedestrian fatalaties increased by 9\% to 5987, representing the highest number since 1990, and also representing 16\% of all automotive fatalities \cite{HighwayTrafficSafetyAdministration2016}. As autonomous vehicles inch closer to widespread adoption, they must have a clear control strategy for pedestrian interaction that can handle a wide variety of pedestrian behaviors while maintaining a reasonable flow of traffic.

@@

The need to further understand pedestrian behavior for autonomous driving has created a growing body of interdisciplinary research. One branch of research has focused on modeling pedestrian behavior given various sensor inputs. Keller et al \cite{Keller2014} presented a study on pedestrian path prediction and action classification (e.g. crossing vs. waiting) using Gaussian process dynamical models and trajectory matching from optical data. Several other techniques for pedestrian trajectory prediction have been proposed, including LQR \cite{Batkovic}, set-based reachability analysis \cite{Koschi2018}, and Markov processes \cite{Karasev2016}

Another branch of literature focuses on pedestrian behavior analysis using empirical data, typically in the form of video analysis. While typically studied for purposes of road design, this body of literature holds promising insights for AV designers. For example, Schroeder and Rouphail \cite{Schroeder2011} explored factors associated with driver yielding behavior at unsignalized pedestrian crossings. Using logistic regression, the authors found that drivers are more likely to yield to assertive pedestrians who walk briskly in their approach to a crosswalk. Kadali and Perumal \cite{RaghuramKadali2012} studied the ``gap acceptance" behavior of pedestrians at mid-block crosswalks through a video graphic survey, and found that the gap accepted for crossing was explained by factors such as crossing direction, vehicle speed, and pedestrian age. Yannis et al. \cite{Yannis2013} and Sun et al.\cite{Sun2002} also found that gap acceptance was influenced by the size of the oncoming vehicle and the presence of other pedestrians.  Lee and Aty \cite{Lee2005} studied interactions in the form of crashes, and found crashes were linked with higher daily traffic.

Finally, a small but rapidly growing body of literature specifically studies the interaction between pedestrians and autonomous vehicles at crosswalks. An excellent review of these studies was conducted by Rasouli et al. \cite{Rasouli}. As an example, Rothenbucher et al. \cite{Rothenbucher2016} studied the interaction between pedestrians and driverless vehicles by constructing a car seat costume to disguise a driver. To improve the issue of trust, several researchers have also developed external communication interfaces to more clearly broadcast the intent of the autonomous vehicle (\cite{Matthews}, \cite{Lagstrom2015}).

One issue with automated driving for crosswalk scenarios is that an overly conservative crossing algorithm will often be taken advantage of or cause confusion among pedestrians\cite{Camara2018}. Camara et al. \cite{Camara2018} attempted to model the natural negotation for priority between a pedestrian and an AV at an intersection using the framework of game theory. Chen et al. \cite{Chen} notes the need for a tradeoff between passive and aggressive driving behavior, and developed a stochastic model of pedestrian behavior to evaluate proposed AV control policies. Control polices for pedestrian interaction are also developed in \cite{Bandyopadhyay}, who proposes a Mixed-observable Markov Decision Process (MOMDP) to incorporate the intention uncertainty of a pedestrian. A Partially-observable Markov Decision Process (POMDP) formulation is also proposed by Thornton \cite{Thornton2018}, who proposes a value-sensitive design framework to assist with the development of engineering specifications for the control algorithm.

@@

This paper aims to address this gap by developing a hybrid control architecture that accounts for several distinct pedestrian modes of behavior at an unsignalized crosswalk. An unsignalized crosswalk is chosen as it is the trickiest crosswalk for a pedestrian and vehicle to navigate, as there is no explicit declaration of whose turn it is cross. Simulations show that the hybrid controller is able to handle a continuous spectrum of pedestrian \textit{gap acceptance} behavior, tolerating a range of highly conservative to highly aggressive pedestrians. Moreover, this is possible with just four distinct controller modes. Finally, experimental results are shown on a real vehicle to demonstrate the feasibility of the proposed controller in real-time.

@@

In general, pedestrian crosswalks may be controlled or uncontrolled. In the case of the former, control devices such as stop lights and walk signals guide the interaction of vehicles and pedestrians explicitly. Fig.~\ref{fig:schematic} shows the latter example, in which there is no such control device, and a pedestrian must select a gap in traffic flow and cross.

\begin{figure}
\centering
\includegraphics[width=3.5in]{figures/example2.eps}
\caption{Top: Description of relevant problem variables. Bottom: Schematic of pedestrian approaching a stream of traffic at a mid-block intersection. Ego vehicle shown in green. We develop a hybrid control framework for the green vehicle, which considers the distance ($d$) to a stopping point $\Delta$ ahead of the crosswalk and the pedestrian's position in the crosswalk ($x_p$).}
\label{fig:schematic}
\end{figure}

In general, right-of-way for uncontrolled intersections is complex. For example, nine states and the District of Columbia require motorists to \textit{stop} when approaching a pedestrian in an uncontrolled crosswalk. Six states require a motorist to stop when a pedestrian is upon the same half of the roadway or within one lane of the lane that the motorist is traveling in. Another nineteen states require a motorist to \textit{yield} when a pedestrian is upon any portion of the roadway, and another 20 states mandate that motorists yield when a pedestrian is upon the same half of the roadway or approaching closely from the opposite side of the roadway \cite{Legislatures2018}.

\subsection{Problem Formalization}
\label{sec:probform}

A diagram of the relevant state variables is shown in Fig.~\ref{fig:schematic}. We assume that knowledge of distance $d$ to a fixed stopping point 3-5 meters ahead of the crosswalk can be estimated by a perception system, and that the velocity of the vehicle $\dot{d}$ is known as well. Furthermore, we also assume the perception system is able to determine the position of the pedestrian $x_p$ within the crosswalk, and is able to determine a velocity estimate $\dot{x}_p$ for the pedestrian as well.

% \begin{figure}
% \centering
% \includegraphics[width=1in]{figures/diagram.eps}
% \caption{}
% \label{fig:diagram1}
% \end{figure}

In general, this control problem is difficult as there are multiple stakeholders and therefore multiple competing objectives that must be considered. For example, relevant stakeholders for this problem are the autonomous vehicle occupants, the pedestrian(s), and the other vehicles in the traffic stream.

As a result, an algorithm that simply yields every time a pedestrian is near the crosswalk is likely to be too conservative, as there is a likelihood the pedestrian is waiting for a larger gap in the traffic stream to cross or is walking through the sidewalk without intending to cross in the first place.

\subsection{System Requirements}

To formalize the system requirements for this multi-stake holder problem, we propose the following \textit{design requirements}, extended from the engineering specifications proposed by Thornton \cite{Thornton2018}.

\begin{center}
\begin{tabular}{ l| l }
\label{tb:specs}
 \textbf{Engineering} & \textbf{Design} \\
 \textbf{Specification\cite{Thornton2018}} & \textbf{Constraint}\\\hline 
 1. Safety and Legality & a. Avoid collisions for all\\
                     & possible accepted gaps                        \\
                     &b. Ability to follow stop/yield    \\
                     &laws for all 50 U.S. states\\

               & \\
 3. Smoothness & a. Low acceleration: Nominal brake \\ 
 			   & and acceleration limits of 2 $\mathrm{m/s^2}$ \\

@@

Similar to time to collision (TTC), the safety gap is a measure of how much time there is before the vehicle would enter the crosswalk if it kept its current speed constant. When faced with a stream of traffic at a crosswalk, a pedestrian inherently decides how much of a gap to accept before crossing. The average gap acceptance is reported in the literature to be between 3-7 seconds, meaning pedestrians usually do not cross if the vehicle would enter the crosswalk in under three seconds \cite{DiPietroCharlesMandKing1970}, and are very likely to cross when they have more than seven seconds \cite{Schmidt2009}.

@@

The proposed control architecture is shown in Fig.~\ref{fig:hybridController}, and the algorithm for operation is shown in Algorithm \ref{alg:hybridController}. At heart, the control algorithm uses a feedback-feedforward methodology to compute desired acceleration commands $\ddot{d}$, but computes these commands differently depending on which of four discrete modes the controller is in. The next section will describe the four modes.

\begin{figure}
\centering
\includegraphics[width=3.5in]{figures/stateDiagram2.eps}
\caption{Schematic of hybrid control structure.}
\label{fig:hybridController}
\end{figure}

\begin{table}[h]
\begin{center}
\caption{Parameters for Algorithm 1}\label{tb:params}
\begin{tabular}{lc}
Symbol & Definition \\\hline\hline
$d_\mathrm{max}$ & $\frac{\dot{d}^2}{2a_\mathrm{max}}$ \\
$d_\mathrm{cmf}$ & $\frac{\dot{d}^2}{2a_\mathrm{cmf}}$ \\\hline
\end{tabular}
\end{center}
\end{table}

@@

		\ElsIf{$d > d_\mathrm{cmf}$}
		\State $state \leftarrow YIELDING$

	% \ElsIf{$d < \frac{\dot{d}^2}{2a_\mathrm{cmf}} \hspace{1mm} \mathbf{and} \hspace{1mm} d > \frac{\dot{d}^2}{2a_\mathrm{max}}$}

  \ElsIf{$d_\mathrm{max} < d < d_\mathrm{cmf} $}

	\State $state \leftarrow HARD\hspace{1mm}BRAKING$

@@

	\EndIf
	\EndIf
	\EndIf
\vspace{3mm}
\If {$state = YIELDING$}
\If {$d > d_\mathrm{cmf} + t_{delay}\dot{d}$}
	\State $\ddot{d}_{des} = 0$
	\State $\dot{d}_{des} = v_\mathrm{speed limit}$
\Else 
	\State $\ddot{d}_{des} = -a_\mathrm{cmf}$
	\State $\dot{d}_{des} = \mathrm{getDesiredSpeed}(d, \dot{d})$
\EndIf 
	\State $\ddot{d} = \ddot{d}_{des} + k_s(\dot{d} - \dot{d}_{des})$

@@

\EndIf
\If {$state = SPEED\hspace{1mm}UP$}
	\State $\ddot{d} = a_\mathrm{cmf}$ 	
	\If {$\textbf{not} \hspace{1mm} \mathrm{inCrosswalk}(x_p) \vee d < 0$} \State $state \leftarrow DRIVING$

@@

\subsection{Nominal Driving}
In the nominal state, the vehicle attempts to drive through the crosswalk at the speed limit $v_\mathrm{speedlimit}$. A simple proportional speed control is applied to keep the vehicle at the speed limit, achieving design constraint 2.a:

@@

The controller remains in the driving mode unless the pedestrian begins to enter the crosswalk before the vehicle has passed. In this case, we mathematically define ``beginning to enter" as the instant the pedestrian begins moving towards the crosswalk, even if the pedestrian is currently on the sidewalk:

\begin{equation}
\label{eq:crossing}
\mathrm{inCrosswalk}(x_p, 
\dot{x}_p) = \left\{
  \begin{array}{ll}
    1&: \dot{x}_p \neq 0 \vee \hspace{1 mm} 0 \leq x_p \leq x_F \\
    0&: \mathrm{otherwise} 
  \end{array}
\right.
\end{equation}

 Where $x_F$ is the end of the crosswalk area of interest. Note that according to design constraint 1.b, the definition of $x_F$ will vary according to which US state the vehicle is in. For example, in California, $x_F$ would be at the end of the crosswalk, while in Louisiana, $x_F$ would just be the same half of the crosswalk as the vehicle \cite{Legislatures2018}.

@@

Frequently the vehicle will pass through the intersection well before the pedestrian will reach the vehicle's lane. Assuming the vehicle is operating in a US state where stopping is not explicitly required, design constraint 2.a suggests it is desirable for the vehicle to continue through the intersection.

This can be formalized by calculating the \textit{time advantage} $t_\mathrm{adv}$, also known as the \textit{time to collision} as follows:

\begin{equation}
t_\mathrm{adv} = \frac{d}{\dot{d}} - \frac{x_v - x_p}{\dot{x_p}}
\end{equation}

Where $x_v$ is the $x$ position of the vehicle in the crosswalk. Algorithm 1 explicitly allows the vehicle to continue driving through the intersection as long as $t_\mathrm{adv}$ exceeds a specified threshold $t_\mathrm{max}$.

@@

If the time advantage is not sufficient for the vehicle to pass through, the autonomous vehicle must slow down. The preferred option to meet design constraint 3.a is to brake at a low, comfortable deceleration $-a_\mathrm{cmf}$ for the passenger.

Comfortable braking at a deceleration $-a_\mathrm{cmf}$ is possible if sufficient braking distance exists. The braking distance is derived kinematically as

\begin{equation}
d_{brake} = \frac{\dot{d}^2}{2a_\mathrm{cmf}}
\end{equation}

If this holds true when the pedestrian starts to enter the crosswalk, the car enters a \textit{Yielding} mode.

@@

In the unlikely event a pedestrian begins to cross when the time gap is low (e.g. under 3 seconds), the vehicle will need to prioritize design constraint 1.a and come to a stop at a deceleration higher than $-a_\mathrm{cmf}$. This occurs when the following conditions holds:

\begin{equation}
\frac{\dot{d}^2}{2a_{max}} < d < \frac{\dot{d}^2}{2a_\mathrm{cmf}}
\end{equation}

@@

\subsubsection{Speeding Up}

As a final condition, consider the case where the pedestrian enters the crosswalk just as the vehicle is crossing $d = 0$. In this case, it makes little sense to decelerate, as the vehicle has insufficient space to stop before the crosswalk and will risk being rear-ended or stopping in the crosswalk. It makes more sense in this condition for the vehicle to speed up and exit the crosswalk quickly. As a result, if $d < \frac{\dot{d}}{2a_{max}}$, the vehicle speeds up.

@@

In the yielding state, the vehicle follows a different version of the feedback-feedforward dynamics. The vehicle first follows the speed limit until it reaches the critical value of $\frac{\dot{d}^2}{2a_\mathrm{cmf}}$. In practice, an additional term $t_\mathrm{delay}\dot{d}$ is added to compensate for the delay in brake communication.

At the critical distance, the vehicle decelerates at the uniform yielding deceleration of $-a_\mathrm{cmf}$, with an additional feedback term $k_s(\dot{d} - \dot{d}_\mathrm{des}$), where $\dot{d}_\mathrm{des}$ is the desired velocity. The feedback term helps bring the vehicle speed to 0 at the desired stopping point.

The desired velocity profile $\dot{d}_\mathrm{des}(d)$ for a vehicle decelerating at constant acceleration is given by:

\begin{equation}
\dot{d}_\mathrm{des}(d) = \sqrt{2a_\mathrm{cmf}(d-d_o) + \dot{d}_o^2}
\end{equation}

Where $d_o$ and $\dot{d}_o$ are the values of $d$ and $\dot{d}$ when the controller first enters the yielding state. Inspection shows that the desired velocity is $\dot{d}_o$ when $d = d_o$ and is 0 when $d = 0$ (recall that we enter the yielding state when $d_o = \frac{\dot{d_o}^2}{2a_\mathrm{cmf}}$).

The controller exits the yielding mode and returns to the driving mode once the pedestrian clears the crosswalk (e.g. once $x > x_F$ in (\ref{eq:crossing})).

@@

In the hard braking state, the vehicle follows a similar feedback-feedforward algorithm, but the desired feedforward deceleration in this case is given by $\ddot{d} = -\frac{\dot{d}^2}{2d}$. Integrating to find the desired speed profile $\dot{d}_\mathrm{des}(d)$ yields the following deceleration:

\begin{equation}
\dot{d}_\mathrm{des}(d) = \frac{\dot{d}_o}{\sqrt{d_o}}\sqrt{d}
\end{equation}

Where $d_o$ and $\dot{d}_o$ are the values of $d$ and $\dot{d}$ when the controller first enters the braking state. Again, the controller exits the braking mode and returns to the driving mode once the pedestrian clears the crosswalk.

@@

In the speed up state, there is no need to follow an exact speed profile as the vehicle is merely trying to exit the crosswalk area slightly faster. In this case, the commanded acceleration is $\ddot{d} = a_\mathrm{cmf}$ until the pedestrian clears the crosswalk or once the vehicle passes the stopping location (i.e. $d< 0$).

@@

In order to check performance of the algorithm, both simulation and experimental studies were undertaken. The simulation results focused on testing Algorithm 1 against the specified design requirements over many pedestrian crossings.

The algorithm is validated in simulation via a built-from-scratch crosswalk simulation environment (Fig. \ref{fig:simFramework}) developed in Python. The simulation environment is built on the ROS robotics interface to enable easy porting of the control architecture to the experimental vehicle in Section \ref{sec:expres}. The simulation includes features such as customizable lanes, the pedestrian crossing on the right or the left, and animation capabilities.

\begin{figure}
\centering
\includegraphics[width=2.5in]{figures/simFramework.png}
\caption{Crosswalk simulation environment for algorithm validation. Ego vehicle is shown in black, leading vehicle in traffic stream shown in blue, and pedestrian shown in red.}
\label{fig:simFramework}
\end{figure}
